Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2021-10-15 doc/AMS-15/AMS-15_poster.Rmd
Add/edit text. Remove graphite value for hgis outlier.

@@

@@

![HGIS Photo](images/gas_control.jpg)
![HGIS Photo](images/gilson.png)

@@

@@

- Errors propagated at each step. 
- Per-sample error agrees well with variability of replicates.

Data reduction done using an R package maintained on github:

https://github.com/blongworth/hgis

@@

Data used for current, blanks, consensus

@@

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```
## Initial measurements

@@

compare to first tests

Note that our flat-topped current curve allows variable flow of the vial system. MICADAS has a more pronounced peak so this may not work as well. (Check papers for peak shape)

Also note current dependent ratio (blank) as a complicating factor

Lead into blank on next slide

@@

#### Current dependent blank

@@

#### Current independent blank

If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.

@@

Data from the measurement of consensus standards over a range of fraction modern shows that this simple correction is effective for removing blank contribution without bias.

#### Mean Fm of blanks: `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`

@@

## Results: Consensus standards

@@

## Results: By sample type

```{r}
ggplot(cons_df, aes(x = fm_diff, y = fm_corr, color = he12C)) +
  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
             color = "grey", linetype = "dashed") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
             color = "grey", linetype = "dashed") +

  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr), 
                  size = .2) +
  labs(subtitle = "Reference and normalization standards",
       y = "Fraction modern (Fm)") +
  facet_wrap(facets = vars(Name), scales = "free") +
  theme_bw() +
  scale_x_continuous(breaks = NULL) +
  xlab(NULL) +
  scale_color_continuous(name = "12C- (μA)") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## Pendleton Haiti Core Carbonates

### Read data, pull graphite results and sample depth, combine

@@

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite.

## Difference plot

@@

```{r}
# No outliers, vs depth
combined_results %>% 
  filter(abs(fm_diff) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Core depth (cm)",
       y = "Fm - mean Fm")
```

```{r}
# No outliers, vs mean_fm
combined_results %>% 
  filter(abs(fm_diff) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Mean Fm",
       y = "Fm - mean Fm")
```

@@

>>>>

![HGIS Photo](images/gas_control.jpg)

Our method uses an open split and glass capillary to control gas flow to the ion source. Compared to closed inlet systems which regulate introduction of a CO2/Helium mixture to the ion source by regulating sample pressure, our method is simpler and faster, but does not maintain a constant flow of CO2 to the source. We compensate for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.

@@

![HGIS Photo](images/gilson.png)

@@

@@

Normalization to standard, blank correction, and estimation of measurement error are done with the following steps:

@@

Errors from normalization and blank correction are propagated at each step. The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.

Data reduction is done using an R package maintained on github (https://github.com/blongworth/hgis).

@@

@@

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```

@@

@@

@@

@@

Data from the measurement of consensus standards over a range of fraction modern shows that this simple correction is effective for removing blank contribution without bias. Our mean measured fraction modern of C-1 and TIRI-F carbonate blanks is `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`.

@@

@@

## Marine Sediment Core Carbonates

To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method. We found the measurements to be in good agreement, with no significant bias and errors averaging XXX compared with XXX for graphite measurements.

@@

@@

```{r}
# No outliers, vs depth
combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Core depth (cm)",
       y = "Fm - mean Fm")
```

```{r, eval = FALSE}
# No outliers, vs mean_fm
combined_results %>% 
  filter(abs(fm_diff_mean) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Mean Fm",
       y = "Fm - mean Fm")
```

@@

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite. This sample was dropped from the comparison. Not sure whether to mention this on poster.
